Diversification of Australian blind snakes

### 2023-01-17 manuscript.Rmd
update abstract section and wrote some manuscript title

---
title: "Diversification of Australian blind snakes"
date: "`r Sys.Date()`"
author: 
output: 
  # bookdown::html_document2:
  #   toc: true
  #   number_sections: FALSE
  bookdown::word_document2:
    reference_docx: [C:/Users/ST/Documents/repo/AusARG_blindsnake/custom-reference-doc.docx]
    number_sections: false
    fig_caption: yes
bibliography: [C:/Users/ST/Documents/repo/anu_thesis_texts/zotero_library.bib]
# biblio-style: "apalike"
csl: jbiogeography.csl
link-citations: true
editor_options:
  chunk_output_type: console
---

@@

@@

@@

**Methods**:

**Results**:

**Main conclusions**:

**Keywords**
Scolecophidia, Diversification, paleoclimate, BioGeoBEARS, GeoHiSSE

@@

We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity (estimated from @scoteseAtlas2021) provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA-TOP10). This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23). The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table-RPANDA-TOP10). Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (see supp-TABLE_FULL_RPANDA_model_fit).

@@

>>>>

---
title: ""
date: "`r Sys.Date()`"
author: 
output: 
  # bookdown::html_document2:
  #   toc: true
  #   number_sections: FALSE
  bookdown::word_document2:
    reference_docx: [C:/Users/ST/Documents/repo/AusARG_blindsnake/custom-reference-doc.docx]
    number_sections: false
    fig_caption: yes
bibliography: [C:/Users/ST/Documents/repo/anu_thesis_texts/zotero_library.bib]
# biblio-style: "apalike"
csl: jbiogeography.csl
link-citations: true
editor_options:
  chunk_output_type: console
---

@@

Possible titles:
- Diversification rates of Australian blind snakes slows down as the continent became more dry up.
- Reconstructed paleoaridity correlate with decreased diversification rates in Australian blind snakes.
- Drying up speciation: aridification of Australia slowed down diversification of Australian blind snakes
- Dull when dry: increased aridity correlate with decrease diversification in blind snakes

@@

**Aim**: Shifts in diversification rates of Australian flora and fauna has been associated with aridification, but the relationship between diversification rates and aridity ifself has never been quantified. We investigate how paleoenvironmental changes have influenced the diversification rates of the mostly endemic Australian blind snakes genus, Anilios. We explored how changes in temperature and aridity during the Neogene have influenced the diversification of the Australian blind snake and we further investigate whether diversification rates differ between arid-adapted lineages vs mesic-adapted lineages.

@@

**Materials and Methods**: We infer the ancestral biome of Australian blind snakes using BioGeoBEARS. We reconstruct the paleoenvironments of Australia during the Neogene using the gen3sis engine. We fit several birth-death models and estimate diversification rates under paleoenvironmental conditions using RPANDA. We further compare diversification rates between arid-adapted lineages versus mesic-adapted lineages using ClaDS and GeoHiSSE.

**Results**: Results from our BioGeoBEARS analyses indicate that Australian blind snakes have tropical grassland origins. We found that local Australian paleoenvironments provide better fit than global paleoenvironments. Diversification rates of blind snakes decreased with increasing aridity. We found no difference in diversification rates between arid- and mesic-adapted lineages.

**Main conclusions**: Soon after colonising Australia, the common ancestors of Australian blind snakes diversified rapidly in mesic habitats during early Miocene. When the continent became more arid, diversification rates decreased. However, two clades emerged with one staying mostly in mesic habitats and the other exploiting the burgeoning arid biome. Both the arid and tropical biomes were important sources and sinks of subsequent diversification.

**Keywords**
Scolecophidia, Diversification, gen3sis, BioGeoBEARS, GeoHiSSE

@@

We found that the exponential environment-dependent speciation and constant extinction (BEnvDCST) model based on paleoaridity (estimated from @scoteseAtlas2021) provided the best fit to the phylogeny (AICc = 238.43; see Table-RPANDA-TOP10). This best-fitting model indicates that an increase in aridity is exponentially associated with a decrease in speciation rate (λ = 1.32, α = -0.14) and a constant background extinction rate (µ = 0.23) (Fig. \@ref(fig:fig-RPANDA)). The second and third most competitive models (ΔAICc < 2) also found that an increase in aridity is exponentially associated with a decrease in speciation rate, but without estimating extinction (see Table-RPANDA-TOP10). Among the exponential environment-dependent models, we found that Australia-specific paleotemperature data provided a better fit to our phylogeny than global paleotemperature data (see supp-TABLE_FULL_RPANDA_model_fit).

```{r fig-RPANDA, fig.cap= 'Effect of aridity change on the diversification of Anilios estimated from the phylogeny. (a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line). (b) Speciation rates through time from the best diversification model. (c) Association between speciation rate and aridity index. We found that increased aridity is associated with decrease speciation rate.'}

knitr::include_graphics(paste0(WD, '/assets/figures/RPANDA_fit_arid.jpg'))
```

@@

Fig. Effect of aridity change on the diversification of Anilios estimated from the phylogeny. (a) The aridity index through time calculated from reconstructed paleoclimate (dotted line), and lineage through time estimated from the phylogeny (solid line). (b) Speciation rates through time from the best diversification model. (c) Association between speciation rate and aridity index. We found that increased aridity is associated with decrease speciation rate.
